\vspace{-1em}Guidance, Please! Towards a Framework for RDF-based Constraint Languages.

### 2015-07-20 paper/paper 2/dc-2015.tex
3.5
formal spec of RDF-CV

\title{\vspace{-1em}Guidance, please! Towards a framework for RDF-based constraint languages.}

@@

We formerly demonstrated that a high-level constraint language like OWL 2 and DSP can be implemented by mapping the language to SPIN using SPARQL CONSTRUCT queries \cite{BoschEckert2014-2}. We provide a validation environment where own mappings from arbitrary constraint languages can be provided and tested\footnote{Online available at: \url{http://purl.org/net/rdfval-demo}, source code online available at: \url{https://github.com/boschthomas/rdf-validator}.}. The only limitations are that the constraints have to
be expressed in RDF and that the constraint language is expressable in SPARQL.

@@

At the core, we use a very simple conceptual model for constraints (see Figure \ref{fig:RDF-CV-conceptual-model}), using a small lightweight vocabulary called \emph{RDF Constraints Vocabulary (RDF-CV)}\footnote{Online available at: \url{https://github.com/boschthomas/RDF-Constraints-Vocabulary}}.\tb{ToDo for Thomas: actualize vocabulary}
\begin{figure}[H]
	\centering
		\includegraphics[width=0.80\textwidth]{images/RDF-CV.png}
	\caption{\emph{RDF Constraints Vocabulary (RDF-CV)} Conceptual Model}
	\label{fig:RDF-CV-conceptual-model}
\end{figure}

@@

\emph{Left} and \emph{right property lists} are lists usually containing properties the constraint is applied to. A typical example for a constraint with a right hand side list of properties would be a constraint of the type \emph{literal value comparison (R-43)} to express that birth dates must be before death dates ({\small\ms{birthDate $<$ deathDate}}). The assignment of properties to these lists happens relative to the constraining element.
\emph{Object property paths} (\emph{R-55}) ensure that if an individual \emph{x} is connected by a sequence of object properties with an individual \emph{y}, 
then \emph{x} is also related to \emph{y} by a particular object property. 
As \emph{Stephen-Hawking} is the author of the book \emph{A-Brief-History-Of-Time} 
%{\small\ms{(authorOf(Stephen-Hawking, A-Brief-History-Of-Time))}} 
whose genre is \emph{Popular-Science}, 
%{\small\ms{(genre(A-Brief-History-Of-Time, Popular-Science)),}}
the object property path {\small\ms{authorOf $\circ$ genre $\sqsubseteq$ authorOfGenre}} infers that \emph{Stephen-Hawking} is an author of the genre \emph{Popular-Science}. 
%{\small\ms{(authorOfGenre(Stephen-Hawking, Popular-Science))}.}
Thus, when representing the constraint using the RDF-CV (see Table \ref{tab:property-constraint-object-property-paths}), the properties \emph{authorOf} and \emph{genre} are placed on the left side of the constraining element \emph{property path}
and the property \emph{authorOfGenre} on its right side. The \emph{context class} limits the constraint to individuals of a specific class. \tb{ToDo: A context class may be an RDFS class, an OWL class, or an RDFS datatype as...} As the \emph{property path} constraint holds for all individuals within the data, the context class is set to the \emph{DL top concept} $\top$ which stands for the super-class of all possible classes.
\begin{table}[H]
  \scriptsize
  \sffamily
  \vspace{0cm}
	\caption{Object Property Paths as Property Constraint}
	\label{tab:property-constraint-object-property-paths}
	\centering
		\begin{tabular}{c|c|c|c|c|c}
      \textbf{context class} & \textbf{left p. list} & \textbf{right p. list} & \textbf{classes} & \textbf{c. element} & \textbf{c. value} \\
      \hline
      $\top$ & authorOf, genre & authorOfGenre & $\top$ & property path & - \\
		\end{tabular}
\end{table}

@@

In this paper, we outlined our idea of a general framework to support the mapping of high-level constraint languages to a generic representation, which subsequently can be used to generate validating SPIN/SPARQL queries to actually validate data against constraints provided in the high-level language.\tb{I do not understand this!} The framework consists of a very simple conceptual model using the \emph{RDF Constraints Vocabulary (RDF-CV)} which has been introduced in this paper. The core of the framework is the definition of 103 constraining elements that are used to define constraints of all 81 constraint types that to date have been identified within the DCMI RDF Application Profiles Task Group and in cooperation with the W3C Data Shapes Working Group.
The full definition of all constraint types and the generic representation of the types in RDF-CV is provided in an accompanying technical report \cite{BoschNolleAcarEckert2015}.

>>>>

\title{\vspace{-1em}Guidance, Please! Towards a Framework for RDF-based Constraint Languages.}

@@

We formerly demonstrated that a high-level constraint language like OWL 2 and DSP can be implemented by mapping the language to SPIN using SPARQL CONSTRUCT queries \cite{BoschEckert2014-2}. We provide a validation environment where own mappings from arbitrary constraint languages can be provided and tested\footnote{Online available at: \url{http://purl.org/net/rdfval-demo}, source code online available at: \url{https://github.com/boschthomas/rdf-validator}.}. The only limitations are that the constraints have to
be expressed in RDF and that the constraint language is expressible in SPARQL.

@@

At the core, we use a very simple conceptual model for constraints (see Figure \ref{fig:RDF-CV-conceptual-model}), using a small lightweight vocabulary called \emph{RDF Constraints Vocabulary (RDF-CV)}\footnote{Formal specification and HTML documentation online available at: \url{https://github.com/boschthomas/RDF-Constraints-Vocabulary}}}.
\begin{figure}[H]
	\centering
		\includegraphics[width=0.80\textwidth]{images/RDF-CV.png}
	\caption{\emph{RDF Constraints Vocabulary (RDF-CV)} Conceptual Model}
	\label{fig:RDF-CV-conceptual-model}
\end{figure}

@@

\emph{Left} and \emph{right property lists} are lists usually containing properties the constraint is applied to. A typical example for a constraint with a right hand side list of properties would be a constraint of the type \emph{literal value comparison (R-43)} to express that birth dates must be before death dates ({\small\ms{birthDate $<$ deathDate}}). The assignment of properties to these lists happens relative to the constraining element.
\emph{Object property paths} (\emph{R-55}) ensure that if an individual \emph{x} is connected by a sequence of object properties with an individual \emph{y}, 
then \emph{x} is also related to \emph{y} by a particular object property. 
As \emph{Stephen-Hawking} is the author of the book \emph{A-Brief-History-Of-Time} 
%{\small\ms{(authorOf(Stephen-Hawking, A-Brief-History-Of-Time))}} 
whose genre is \emph{Popular-Science}, 
%{\small\ms{(genre(A-Brief-History-Of-Time, Popular-Science)),}}
the object property path {\small\ms{authorOf $\circ$ genre $\sqsubseteq$ authorOfGenre}} infers that \emph{Stephen-Hawking} is an author of the genre \emph{Popular-Science}. 
%{\small\ms{(authorOfGenre(Stephen-Hawking, Popular-Science))}.}
Thus, when representing the constraint using the RDF-CV (see Table \ref{tab:property-constraint-object-property-paths}), the properties \emph{authorOf} and \emph{genre} are placed on the left side of the constraining element \emph{property path}
and the property \emph{authorOfGenre} on its right side. The \emph{context class} limits the constraint to individuals of a specific class. A context class may be an rdfs:Class, an owl:Class (as sub-class of rdfs:Class), or an rdfs:Datatype which is both an instance of and a sub-class of rdfs:Class. As the \emph{property path} constraint holds for all individuals within the data, the context class is set to the \emph{DL top concept} $\top$ which stands for the super-class of all possible classes.
\begin{table}[H]
  \scriptsize
  \sffamily
  \vspace{0cm}
	\caption{Object Property Paths as Property Constraint}
	\label{tab:property-constraint-object-property-paths}
	\centering
		\begin{tabular}{c|c|c|c|c|c}
      \textbf{context class} & \textbf{left p. list} & \textbf{right p. list} & \textbf{classes} & \textbf{c. element} & \textbf{c. value} \\
      \hline
      $\top$ & authorOf, genre & authorOfGenre & $\top$ & property path & - \\
		\end{tabular}
\end{table}

@@

In this paper, we outlined our idea of a general framework to support the mapping of high-level constraint languages to a generic representation, which subsequently can be used to generate validating SPIN/SPARQL queries to actually validate data against constraints provided in the high-level language.\tb{Aus der generischen Repräsentierung können die SPIN mappings zur Validierung generiert werden? das verstehe ich nicht} The framework consists of a very simple conceptual model using the \emph{RDF Constraints Vocabulary (RDF-CV)} which has been introduced in this paper. The core of the framework is the definition of 103 constraining elements that are used to define constraints of all 81 constraint types that to date have been identified within the DCMI RDF Application Profiles Task Group and in cooperation with the W3C Data Shapes Working Group.
The full definition of all constraint types and the generic representation of the types in RDF-CV is provided in an accompanying technical report \cite{BoschNolleAcarEckert2015}.
